
```{r}
library(caTools)
setwd("F:/Study Material/Semester 5/Regression and Time Series Model/RTSM_Project")
hour = read.csv("hour.csv", stringsAsFactors = FALSE)

```

Creating a dataframe which consist of day wise entries.
```{r}
library(dplyr)
by_user = group_by(hour, dteday)
names(hour)
day  = summarize(by_user, 
                           temp = mean(temp),
                          atemp = mean(atemp),
                          hum = mean(hum),
                          windspeed = mean(windspeed),
                 casual = sum(casual),
                 registered = sum(registered),
                 cnt = sum(cnt),
                 season = (mean(season)),
                 holiday = (mean(holiday)),
                 weekday = (mean(weekday)),
                 mnth = (mean(mnth)),
                 yr = (mean(yr)),
                 weathersit = round(mean(weathersit),0),
                 workingday = mean(workingday))
summary(day)


```

Changing the categorical variables into factor datatype so that the dummy variables can be generated while performing regression analysis.
```{r}
str(hour)
summary(hour)
hour$season = as.factor(hour$season)
hour$hr = as.factor(hour$hr)
hour$yr = as.factor(hour$yr)
hour$holiday = as.factor(hour$holiday)
hour$weekday = as.factor(hour$weekday)
hour$workingday = as.factor(hour$workingday)
hour$mnth = as.factor(hour$mnth)
hour$weathersit = as.factor(hour$weathersit)

day$season = as.factor(day$season)
day$yr = as.factor(day$yr)
day$holiday = as.factor(day$holiday)
day$weekday = as.factor(day$weekday)
day$workingday = as.factor(day$workingday)
day$mnth = as.factor(day$mnth)
day$weathersit = as.factor(day$weathersit)
summary(day)
``{r}

```


Now we will check the skewness of outcome variables. The skewness is calculated by the skewness function in the package "e1071". The skewness for the cnt variable i.e the outcome variable for hourly data came out to be 1.277191, which tells us that data is highly positively skewed. So, inorder to minimize the skewness and dispersion of the data, logarithmic transformation and normalization are performed on the outcome variable of hourly data. 
On the other hand the daily data had a skewness of -0.04715862. Which is an evidence of weak skewness on the negative side. and hence only normalization is performed in this case.

```{r}
library(e1071)
skewness(hour$cnt)
skewness(day$cnt)

hist(((hour$cnt)),breaks = 10)
hist(day$cnt)

hour$log_cnt = log(hour$cnt)
hour$norm_log_cnt = (hour$log_cnt - min(hour$log_cnt))/(max(hour$log_cnt)-min(hour$log_cnt))
day$norm_cnt = (day$cnt - min(day$cnt))/(max(day$cnt) - min(day$cnt))
plot(hour$temp,hour$atemp)

```

Multiple linear regression model is trained over the hourly data as well as daily data. First of all I have considered a model consisting of all the variables as independent variables except the outcome variable cnt. After checking the correlation matrix we found out some of the variables were highly correlated and hence one of them is considered into the model in order to improve the quality of the model.
Finally I used the step function available in R and found the best adjusted R square value of 0.8235.

```{r}
relation = norm_log_cnt ~ season + yr + mnth + hr + holiday + weekday + workingday + weathersit + temp + atemp + hum + windspeed


model_0 = lm(relation, data = hour[1:nrow(hour),])
summary(model_0)
model = step(model_0)
summary(model)

relation_3 = norm_cnt ~ season + yr + mnth + holiday + weekday + weathersit+ workingday + temp + atemp + hum + windspeed
model_3 = lm(relation_3, data = day)
summary(model_3)
model_day = step(model_3)
summary(model_day)
```

Performing the Residual Analysis 


```{r}

library(stdres)
hist(model$residuals, breaks = 100)
hist(model_day$residuals, breaks = 100)

plot(predict(model), model_1$residuals)
plot(predict(model_day), model_day$residuals)

qqnorm(model$residuals)
qqline(model$residuals)

qqnorm(model_day$residuals)
qqline(model_day$residuals)

```

Performing Time Series Analysis for day-wise data
```{r}
library('ggplot2')
library('forecast')
library('tseries')


# Converting Dates to the Date datatype
day$Date = as.Date(day$dteday)

# plotting the Daily Bike checkouts with date
ggplot(day, aes(Date, cnt)) + geom_line() + scale_x_date('month') + ylab('Daily Bike Checkouts') +
  xlab("")


```

```{r}
# removing timeseries outliers from the data
count_ts = ts(day[,c('cnt')])
day$clean_cnt = tsclean(count_ts)
ggplot() +geom_line(data = day, aes(x = Date,y = clean_cnt))+ ylab('Clean Bike Count')

```

```{r}
# fitting an imaginary line through the dataset using moving averages
# Weekly MA
day$cnt_ma = ma(day$clean_cnt, order = 7)
# Monthly MA
day$cnt_ma30 = ma(day$clean_cnt, order = 30)
#plotting the line 
ggplot() + 
 geom_line(data = day, aes(x = Date, y = cnt_ma, color = "Weekly moving average")) +
 geom_line(data = day, aes(x = Date, y = cnt_ma30, color = "Monthly Moving average")) +
 ylab("Bicycle count")
```

```{r}
#Decomposition of the Data

  #storing the smoothed dataset in a timeseries 
count_ma = ts(na.omit(day$cnt_ma), frequency = 30)
  # Decomposing the timeseries
decomp = stl(count_ma, s.window = 'periodic')
  # deseasonalizing the time series data
deseasonal_cnt <- seasadj(decomp)
plot(deseasonal_cnt)
plot(decomp)

```

```{r}
# Stationarity Test
# Performing Augmented Dickey Fuller Test
adf.test(count_ma, alternative = "stationary")

#Plotting ACF and PACF plots
Acf(count_ma,main = '')
Pacf(count_ma,main="")

#differencing
count_d1 = diff(deseasonal_cnt, differences = 1)
plot(count_d1)
adf.test(count_d1, alternative = "stationary")

```

```{r}
#Using Auto Arima to determine the orders of time series
auto.arima(deseasonal_cnt, seasonal = FALSE)

#Evaluation and Iteration
fit = auto.arima(deseasonal_cnt, seasonal = FALSE)
tsdisplay(residuals(fit), lag.max = 45, main = '(1,1,1) Model Residuals')

fit_2 = arima(x = deseasonal_cnt, order = c(1,1,7))
tsdisplay(residuals(fit_2), lag.max = 45, main = '(1,1,7) Model Residuals')
fit_2

```

```{r}
#forecasting using the fitted model
fcast = forecast(fit_2, h = 30)
fcast
plot(fcast)
i = 0
sse = 0
for (i in range(1:30)) {
    sse = sse + power((fcast$mean[i]-deseasonal_cnt[i]),2)
}
```

